---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Body Characteristics of Dragon Species Evolved in Different Continents {#morph}

```{r load1, echo = FALSE, eval = TRUE, message = FALSE}
library(DBI)
library(tidyverse)

dragons_db <- dbConnect(RSQLite::SQLite(), "../../Course Material/Data/dragons/dragons.db")

dragons <- dbGetQuery(dragons_db, "SELECT * FROM dragons;")
capture_sites <- dbGetQuery(dragons_db, "SELECT * FROM capture_sites;")
captures <- dbGetQuery(dragons_db, "SELECT * FROM captures;")
morphometrics <- dbGetQuery(dragons_db, "SELECT * FROM morphometrics;")
diet <- dbGetQuery(dragons_db, "SELECT * FROM diet;")
tags <- dbGetQuery(dragons_db, "SELECT * FROM tags;")
deployments <- dbGetQuery(dragons_db, "SELECT * FROM deployments;")
telemetry <- dbGetQuery(dragons_db, "SELECT * FROM gps_data;")
```

Dragons found today in Utah belong to 10 different species that evolved in 
different parts of the world. 

```{r}
dragons %>% 
  pull(species) %>% 
  unique()
```

Previous literature has shown that closeness of geographic origin corresponds to
taxonomic relatedness. Based on this, we hypothesized that dragons that evolved 
in the same continent would be more similar in morphometrics characteristics 
than dragons that evolved in different continents, because they are taxonomically
more distant. Many of the species that today live in Utah evolved in Europe. We
want to compare these dragons to those who evolved in South America, Asia, and
Oceania, even though we only have one representative species for each of those 
continents. We expect European dragon species to be more similar between each
other than to other dragon species in body size. 

Let's reclassify dragons based on the continent where they evolved:

```{r}

dragons <- dragons %>% 
  mutate(continent = case_when(
    species %in% c("Hebridean Black",
                   "Romanian Longhorn",
                   "Ukrainian Ironbelly",
                   "Norwegian Ridgeback",
                   "Common Welsh Green",
                   "Swedish Short-Snout",
                   "Hungarian Horntail") ~ "Europe",
    species == "Peruvian Vipertooth" ~ "South America",
    species == "Chinese Fireball" ~ "Asia",
    species == "Antipodean Opaleye" ~ "Oceania"
  ))

```

Now let's plot mean body size of adult dragons by species and continent of origin:

```{r}

reordered_levels <- dragons %>% 
  select(species, continent) %>% 
  distinct() %>% 
  arrange(continent) %>% 
  pull(species)

morphometrics %>% 
  left_join(dragons, by = "dragon_id") %>% 
  filter(age_class == "Adult") %>% 
  group_by(species, continent) %>% 
  summarize(mean_body_size = mean(total_body_length_cm, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(species = factor(species, levels = reordered_levels)) %>% 
  ggplot(aes(x = species, y = mean_body_size, fill = continent)) +
  geom_bar(stat = "identity") +
  labs(x = " ", y = "Mean body size (cm)", fill = "Continent") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))

```

Let's do the same thing with wingspan:

```{r}

morphometrics %>% 
  left_join(dragons, by = "dragon_id") %>% 
  filter(age_class == "Adult") %>% 
  group_by(species, continent) %>% 
  summarize(mean_wingspan = mean(wingspan_cm, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(species = factor(species, levels = reordered_levels)) %>% 
  ggplot(aes(x = species, y = mean_wingspan, fill = continent)) +
  geom_bar(stat = "identity") +
  labs(x = " ", y = "Mean wingspan (cm)", fill = "Continent") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))

```

It looks like our hypothesis was wrong. Dragons from Europe do not seem to differ
more from dragons from other continents than they do between each other.