---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Diet Composition of Adult Dragons in Utah {#diet-comp}

```{r load2, echo = FALSE, eval = TRUE, message = FALSE}
library(DBI)

dragons_db <- dbConnect(RSQLite::SQLite(), "../../Course Material/Data/dragons/dragons.db")

dragons <- dbGetQuery(dragons_db, "SELECT * FROM dragons;")
capture_sites <- dbGetQuery(dragons_db, "SELECT * FROM capture_sites;")
captures <- dbGetQuery(dragons_db, "SELECT * FROM captures;")
morphometrics <- dbGetQuery(dragons_db, "SELECT * FROM morphometrics;")
diet <- dbGetQuery(dragons_db, "SELECT * FROM diet;")
tags <- dbGetQuery(dragons_db, "SELECT * FROM tags;")
deployments <- dbGetQuery(dragons_db, "SELECT * FROM deployments;")
telemetry <- dbGetQuery(dragons_db, "SELECT * FROM gps_data;")
```

Dragons have been documented to prey on livestock in Utah, but so far no data 
existed to demonstrate whether dragons also prey on other animals and in what 
proportion. Here, we look at diet composition of adult dragons in Utah, and we
explore the correlation between the proportion of livestock in the diet and the
body size adult dragons reach. 

```{r}

diet %>% 
  left_join(dragons, by = "dragon_id") %>% 
  filter(age_class == "Adult") %>% 
  group_by(item) %>% 
  tally() %>% 
  mutate(prop = n/nrow(diet)) %>% 
  ggplot(aes(x = item, y = prop)) + 
  geom_bar(stat = "identity") +
  labs(x = " ", y = "Proportion", fill = "Type") +
  theme_bw()
  
```

While domestic goats are the main diet item found in dragon diet samples, followed
by domestic cows, dragon diet also includes wild animals such as mule deer, 
mountain goat, moose, and (rarely) even coyote and cougar. 

Let's explore how livestock consumption relates with body measurements at maturity:

```{r}

totals <- diet %>% 
  group_by(dragon_id) %>% 
  tally() %>% 
  rename(tot = n)

diet %>% 
  mutate(livestock = case_when(
    item %in% c("Cougar", "Coyote", "Moose", "Mountain goat", "Mule deer") ~ FALSE,
    TRUE ~ TRUE
  )) %>% 
  group_by(dragon_id, livestock) %>% 
  tally() %>% 
  left_join(totals, by = "dragon_id") %>% 
  filter(livestock) %>% 
  mutate(prop_livestock = n/tot) %>% 
  left_join(dragons, by = "dragon_id") %>% 
  filter(age_class == "Adult") %>% 
  left_join(morphometrics, by = "dragon_id") %>% 
  ggplot(aes(x = prop_livestock, y = total_body_length_cm)) + 
  geom_point() + 
  labs(x = "Proportion of livestock in diet", 
       y = "Total body length (cm)") +
  theme_bw()

```

It looks like high livestock consumption correlates with increased body size in 
adult dragons. 